---
title: "Global Value Chain Decomposition with decompr"
author: "Sebastian Krantz"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{decompr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width = 7
)
options(width = 90)
```

**decompr** implements four gross-export decompositions from the Global Value Chain (GVC)
literature. All decompositions operate on an inter-country input-output (ICIO) table and
answer the question: *where does the value added embodied in a country's exports
ultimately originate?*

| Function | Reference | Level | Terms |
|----------|-----------|-------|------:|
| `leontief()` | Hummels, Ishii & Yi (2001) | country-industry origin | continuous shares |
| `kww()` | Koopman, Wang & Wei (2014) | country | 9 |
| `wwz()` | Wang, Wei & Zhu (2013) | bilateral country-industry | 16 |
| `bm()` | Borin & Mancini (2019) | country / sector / bilateral | up to 13 |

`bm()` is the recommended state-of-the-art decomposition. Its world/sink perspective also
provides a corrected version of the biased KWW decomposition.

## The leather dataset

The package ships with `leather`, a minimal 3-country × 3-industry ICIO table covering
the leather GVC for Argentina, Turkey, and Germany.

```{r data}
library(decompr)
data(leather)

leather$countries
leather$industries

# Intermediate demand matrix (9 × 9)
leather$inter

# Final demand matrix (9 × 3)
leather$final

# Gross output vector (length 9)
leather$output
```

The rows and columns of `inter` and the rows of `final` are ordered by country first,
then industry, yielding 9 country-industry combinations.

## Building a `decompr` object

`load_tables_vectors()` parses the raw ICIO matrices and pre-computes all quantities
needed for the decompositions: the Leontief inverse `B`, its domestic (`Bd`) and
foreign (`Bm`) parts, the domestic Leontief inverse `L`, value-added coefficients
`Vc`, and export vectors `E`, `ESR`, `Eint`, `Efd`.

```{r load}
x <- load_tables_vectors(leather)
class(x)
names(x)
```

The key scalars:

```{r dims}
x$G  # number of countries
x$N  # number of industries
```

The value-added coefficient vector `Vc = v / o` gives the direct value-added share
of each country-industry's gross output:

```{r vc}
round(x$Vc, 3)
```

Total exports by country-industry:

```{r exports}
round(x$E, 2)
```

## Leontief decomposition

`leontief()` pre-multiplies the Leontief inverse by the value-added coefficient matrix
and post-multiplies by exports, yielding the value-added origin (rows) of each
exporting country-industry (columns).

```{r leontief}
leo <- leontief(x)
head(leo, 12)
```

Each row identifies a *source* country-industry, a *using* country-industry, and the
amount of value added from the source embodied in the using sector's exports. Setting
`long = FALSE` returns the underlying matrix directly.

## Koopman-Wang-Wei (KWW) decomposition

`kww()` aggregates to the country level and splits exports into 9 components covering
domestic value added (DVA), foreign value added (FVA), and various double-counting terms.

```{r kww}
kww(x)
```

> **Note:** The KWW decomposition contains a known systematic bias — it
> underestimates foreign value added by conflating some of it with domestic
> double-counting. Use `bm(perspective = "world", approach = "sink")` for the
> Borin-Mancini correction (see below).

## Wang-Wei-Zhu (WWZ) decomposition

`wwz()` operates at the bilateral country-industry level and decomposes exports into
16 value-added and double-counting terms by importing country.

```{r wwz}
wz <- wwz(x)
dim(wz)
dimnames(wz)
```

The result is a 3-dimensional array: exporting country-industry × term × importing
country. The last four terms (`texp`, `texpint`, `texpfd`, `texpdiff`) are
accounting diagnostics rather than decomposition components.

```{r wwz-slice}
# DVA absorbed by final demand in Germany (importer index 3)
round(wz[, "DVA_FIN", "Germany"], 2)
```

`wwz2kww()` maps the 16-term result to the 9-term KWW format when both are needed:

```{r wwz2kww}
wwz2kww(wz)
```

### WWZ vs. BM (exporter/source): different perimeters

Although both WWZ and BM decompose bilateral exports into value-added and double-counting
components, they differ fundamentally in *what they define as double-counted*, and this
shapes which empirical questions each can answer.

**WWZ** defines double-counting at the *bilateral* level: an item is called double-counted
only if it crosses the same bilateral border s→r more than once. This is a less
restrictive notion — some items that would be double-counted from an exporting-country
perspective are treated as value-added here. The bilateral perspective answers the question
*"how much of any country's GDP is embedded in this specific trade flow?"*, which is
useful for assessing the GDP-at-stake from a bilateral policy shock. However, because the
perimeter shifts with every (s, r) pair, WWZ results are *not additive*: summing bilateral
terms over all importers r does not recover a meaningful country-level decomposition. In
addition, WWZ mixes source-based and sink-based approaches in the same decomposition —
different terms reflect different accounting logics, making them not directly comparable
and preventing the construction of consistent GVC participation indicators such as DAVAX.

**BM (exporter/source)** defines double-counting at the *exporting-country* level: an item
is double-counted the second time it crosses country s's border, regardless of destination.
This *country-level perspective* is internally consistent (one approach applied throughout)
and fully additive — bilateral results sum to sector totals, sector totals sum to country
totals. It directly yields DAVAX (domestic value added directly absorbed by the importer),
GVC-related trade, and backward/forward GVC participation shares. For most country- or
sector-level analysis BM (exporter/source) is the appropriate choice; WWZ is better suited
when the focus is on the GDP content of a *specific bilateral flow* as such (e.g. tariff
incidence analysis, bilateral trade balance decomposition).

## Borin-Mancini (BM) decomposition

`bm()` is the most flexible decomposition. It supports three aggregation levels
(`"country"`, `"sector"`, `"bilateral"`) and two accounting perspectives
(`"exporter"` with 13 terms, or `"world"` with 9 terms).

### Country level (exporter perspective, 13 terms)

```{r bm-country}
bm(x)
```

The 13 terms break gross exports (`GEXP`) into domestic content (`DC`) and foreign
content (`FC`), and further into value-added and double-counting components. GVC
participation is measured by `GVC = GEXP - DAVAX`, split into backward (`GVCB`) and
forward (`GVCF`) linkages.

### Sector level

```{r bm-sector}
bm(x, aggregation = "sector")
```

### Bilateral sector level

```{r bm-bilateral}
bm(x, aggregation = "bilateral")
```

The bilateral result contains one row per (exporting country-industry, importing
country) pair, excluding within-country flows. Country- and sector-level results are
additive aggregations of this table.

### Corrected KWW (world perspective, 9 terms)

The world/sink perspective implements the Borin-Mancini correction to the biased KWW
decomposition. It is only available at the country level.

```{r bm-worldsink}
bm(x, perspective = "world", approach = "sink")
```

Comparing to `kww()` above, `FVA` increases and `DDC` decreases accordingly.

## Convenience interface

`decomp()` provides a single entry point for all methods. It accepts either a
`decompr` object or a raw `iot` list directly:

```{r decomp}
# Pass the raw iot list — load_tables_vectors() is called internally
decomp(leather, method = "leontief")
decomp(leather, method = "bm")
```

## References

Borin, A., & Mancini, M. (2019). Measuring what matters in global value chains and
value-added trade. *World Bank Policy Research Working Paper 8804*.

Koopman, R., Wang, Z., & Wei, S.-J. (2014). Tracing value-added and double counting
in gross exports. *American Economic Review, 104*(2), 459–494.

Wang, Z., Wei, S.-J., & Zhu, K. (2013). *Quantifying international production sharing
at the bilateral and sector levels*. NBER Working Paper 19677.

Hummels, D., Ishii, J., & Yi, K.-M. (2001). The nature and growth of vertical
specialization in world trade. *Journal of International Economics, 54*(1), 75–96.
